Assessing Gender Bias in Machine Translation -- A Case Study with Google Translate

### 2018-03-14 Paper/main.tex
Fixed bug in stats

% Table with our gender bias results grouped by category
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)	& Male ($\%$)	& Neutral ($\%$) \\ \hline
	\hline		
	Education / Training / Library 			& $25.207$ 		& $59.917$		& $2.893$		\\ \hline
	Business / Financial 					& $6.324$ 		& $67.194$		& $7.312$		\\ \hline
	Office / Administrative					& $10.031$		& $62.382$		& $7.732$		\\ \hline
	Healthcare support						& $29.545$		& $51.705$		& $7.955$		\\ \hline
	Management								& $9.486$		& $68.182$		& $3.953$		\\ \hline
	Installation / Maintenance / Repair		& $3.596$		& $68.332$		& $11.588$		\\ \hline
	Healthcare practitioners / technical	& $21.564$		& $61.734$		& $6.977$		\\ \hline
	Community and social service			& $15.584$ 		& $63.636$		& $4.545$		\\ \hline
	Sales and related						& $8.117$ 		& $69.156$		& $8.766$		\\ \hline
	Production								& $14.36$ 		& $53.306$		& $13.120$		\\ \hline
	Architecture and engineering			& $2.194$		& $79.624$		& $5.016$		\\ \hline
	Life, physical, and social sciences		& $3.743$ 		& $82.888$		& $3.743$		\\ \hline
	Transportation / material moving		& $7.532$ 		& $66.234$		& $11.688$		\\ \hline
	Arts / Entertainment					& $7.862$ 		& $71.007$		& $4.423$		\\ \hline
	Legal									& $11.688$ 		& $67.532$		& $5.195$		\\ \hline
	Protective service						& $6.818$ 		& $75.000$		& $5.844$		\\ \hline
	Food prepaparation / serving			& $16.578$ 		& $59.893$		& $10.160$		\\ \hline
	Farming, fishing, and forestry			& $10.490$ 		& $62.238$		& $8.392$		\\ \hline
	Computer and mathematical				& $2.841$ 		& $76.136$		& $4.545$		\\ \hline
	Personal care and service				& $20.937$ 		& $55.096$		& $9.091$		\\ \hline
	Construction and extraction				& $9.492$ 		& $61.631$		& $11.096$		\\ \hline
	Cleaning / Maintenance					& $11.818$		& $63.636$		& $8.182$		\\ \hline
	\hline
	Total           						& $11.107$ 		& $63.030$ 		& $9.278$ 		\\ \hline
	\end{tabular}
	\caption{Percentage of female, male and neutral gender pronouns obtained for each occupation category, averaged over all occupations in said category and tested languages detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-category}
	}
\end{table}

% Table with our gender bias results grouped by category (but categories are further grouped into "super-categories")
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)			& Male ($\%$)			& Neutral ($\%$)	\\ \hline
	\hline
	Education								& $25.207$				& $59.917$				& $2.893$			\\ \hline
	Corporate								& $7.955$				& $68.030$				& $6.364$			\\ \hline
	Service									& $9.221$				& $64.987$				& $9.506$			\\ \hline
	Healthcare								& $23.729$				& $59.014$				& $7.242$			\\ \hline
	Production								& $14.360$				& $53.306$				& $13.120$			\\ \hline
	STEM									& $2.992$				& $80.322$				& $4.373$			\\ \hline
	Arts / Entertainment					& $7.862$				& $71.007$				& $4.423$			\\ \hline
	Legal									& $11.688$				& $67.532$				& $5.195$			\\ \hline
	Farming / Fishing / Forestry			& $10.490$				& $62.238$				& $8.392$			\\ \hline
	Construction / Extraction				& $9.492$				& $61.631$				& $11.096$			\\ \hline
	\hline
	Total           						& $11.107$ 				& $63.030$ 				& $9.278$ 			\\ \hline
	\end{tabular}
	\caption{Percentage of female, male and neutral gender pronouns obtained for each occupation category, averaged over all occupations in said category and tested languages detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-category-grouped}
	}
\end{table}

@@

% Table with our gender bias results grouped by language
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Language 	& Female ($\%$) 	& Male ($\%$)		& Neutral ($\%$)	\\ \hline
	\hline
	Malay     	& $4.122$ 			& $95.780$  		& $0.000$			\\ \hline
	Estonian  	& $18.351$			& $80.962$  		& $0.491$			\\ \hline
	Finnish   	& $33.562$			& $66.143$  		& $0.000$			\\ \hline
	Hungarian 	& $36.016$			& $63.297$  		& $0.393$			\\ \hline
	Armenian  	& $9.814$			& $81.256$  		& $0.589$			\\ \hline
	Japanese  	& $0.196$			& $26.791$  		& $1.079$			\\ \hline
	Turkish   	& $2.355$			& $63.199$  		& $18.449$			\\ \hline
	Yoruba    	& $1.178$			& $48.577$  		& $36.997$			\\ \hline
	Basque    	& $0.393$			& $12.365$  		& $34.838$			\\ \hline
	Swahili   	& $14.328$ 			& $85.574$  		& $0.000$			\\ \hline
	Chinese   	& $1.865$ 			& $69.382$  		& $9.225$			\\ \hline \hline
	Total     	& ? 				& ? 				& ? 				\\ \hline
	\end{tabular}
	\caption{Percentage of female, male and neutral gender pronouns obtained for each language, averaged over all occupations detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-language}
	}
\end{table}

@@

% Table with our gender bias results for each adjective
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	\hline
	Adjective 	& Female 	& Male 	& Neutral 	& Total 	\\ \hline
	\hline
	Shy     	& 6  		& 2   	& 2  		& 12		\\ \hline
	Attractive 	& 4  		& 2   	& 4  		& 12		\\ \hline
	Happy   	& 5  		& 3   	& 2  		& 12		\\ \hline
	Kind    	& 4  		& 3   	& 1  		& 12		\\ \hline
	Ashamed 	& 4  		& 5   	& 1  		& 12		\\ \hline
	Smart   	& 2  		& 5   	& 3  		& 12		\\ \hline
	Envious 	& 2  		& 6   	& 1  		& 12		\\ \hline
	Sad     	& 2  		& 6   	& 2  		& 12		\\ \hline
	Loving  	& 2  		& 6   	& 2  		& 12		\\ \hline
	Helpless 	& 2  		& 6   	& 2  		& 12		\\ \hline
	Brave   	& 2  		& 7   	& 1  		& 12		\\ \hline
	Proud   	& 2  		& 7   	& 1  		& 12		\\ \hline
	Hateful 	& 1  		& 5   	& 3  		& 12		\\ \hline
	Dumb    	& 1  		& 6   	& 2  		& 12		\\ \hline
	Innocent 	& 1  		& 8   	& 1  		& 12		\\ \hline
	Right   	& 0  		& 7   	& 3  		& 12		\\ \hline
	Guilty 		& 0 		& 9		& 0			& 12		\\ \hline \hline	
	Total   	& 43 		& 129 	& 41 		& 264		\\ \hline
	\end{tabular}
	\caption{Number of female, male and neutral pronominal genders in the translated sentences for each selected adjective.}
	\label{tab:gender-by-adjective}
	}
\end{table}

>>>>

% Table with our gender bias results grouped by category
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)	& Male ($\%$)	& Neutral ($\%$) \\ \hline
	\hline		
	Education / Training / Library 			& $23.737$ 		& $58.081$		& $3.535$		\\ \hline
	Business / Financial 					& $6.039$		& $62.319$		& $8.937$		\\ \hline
	Office / Administrative					& $9.962$		& $56.322$		& $9.451$		\\ \hline
	Healthcare support						& $27.083$		& $50.000$		& $9.722$		\\ \hline
	Management								& $11.594$		& $61.111$		& $4.831$		\\ \hline
	Installation / Maintenance / Repair		& $3.297$		& $62.393$		& $14.164$		\\ \hline
	Healthcare practitioners / technical	& $18.088$		& $61.499$		& $8.527$		\\ \hline
	Community and social service			& $14.286$ 		& $60.317$		& $5.556$		\\ \hline
	Sales and related						& $7.540$		& $64.683$		& $10.714$		\\ \hline
	Production								& $15.32$		& $45.370$		& $15.825$		\\ \hline
	Architecture and engineering			& $1.916$		& $75.862$		& $6.130$		\\ \hline
	Life, physical, and social sciences		& $3.595$		& $80.065$		& $4.575$		\\ \hline
	Transportation / material moving		& $7.778$		& $60.159$		& $14.286$		\\ \hline
	Arts / Entertainment					& $7.808$		& $66.366$		& $5.405$		\\ \hline
	Legal									& $9.524$ 		& $65.079$		& $6.349$		\\ \hline
	Protective service						& $6.746$		& $71.032$		& $7.143$		\\ \hline
	Food prepaparation / serving			& $14.379$ 		& $56.863$		& $12.418$		\\ \hline
	Farming, fishing, and forestry			& $12.821$ 		& $53.846$		& $10.256$		\\ \hline
	Computer and mathematical				& $3.472$		& $70.833$		& $5.556$		\\ \hline
	Personal care and service				& $17.845$ 		& $52.862$		& $11.111$		\\ \hline
	Construction and extraction				& $9.967$		& $54.739$		& $13.562$		\\ \hline
	Cleaning / Maintenance					& $12.222$		& $57.778$		& $10.000$		\\ \hline
	\hline
	Total           						& $11.078$ 		& $57.398$ 		& $11.286$ 		\\ \hline
	\end{tabular}
	\caption{Percentage of female, male and neutral gender pronouns obtained for each occupation category, averaged over all occupations in said category and tested languages detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-category}
	}
\end{table}

% Table with our gender bias results grouped by category (but categories are further grouped into "super-categories")
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Category 								& Female ($\%$)			& Male ($\%$)			& Neutral ($\%$)	\\ \hline
	\hline
	Education								& $23.737$				& $58.081$				& $3.535$			\\ \hline
	Corporate								& $8.519$				& $62.407$				& $7.778$			\\ \hline
	Service									& $8.730$				& $59.746$				& $11.619$			\\ \hline
	Healthcare								& $20.527$				& $58.380$				& $8.851$			\\ \hline
	Production								& $15.32$				& $45.370$				& $15.825$			\\ \hline
	STEM									& $2.954$				& $76.653$				& $5.345$			\\ \hline
	Arts / Entertainment					& $7.808$				& $66.366$				& $5.405$			\\ \hline
	Legal									& $9.524$				& $65.079$				& $6.349$			\\ \hline
	Farming / Fishing / Forestry			& $12.821$				& $53.846$				& $10.256$			\\ \hline
	Construction / Extraction				& $9.967$				& $54.739$				& $13.562$			\\ \hline
	\hline
	Total           						& $11.078$ 				& $57.398$ 				& $11.286$ 			\\ \hline
	\end{tabular}
	\caption{Percentage of female, male and neutral gender pronouns obtained for each occupation category, averaged over all occupations in said category and tested languages detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-category-grouped}
	}
\end{table}

@@

% Table with our gender bias results grouped by language
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|}
	\hline
	Language 	& Female ($\%$) 	& Male ($\%$)		& Neutral ($\%$)	\\ \hline
	\hline
	Malay     	& $4.122$ 			& $95.780$  		& $0.000$			\\ \hline
	Estonian  	& $18.351$			& $80.962$  		& $0.491$			\\ \hline
	Finnish   	& $33.562$			& $66.143$  		& $0.000$			\\ \hline
	Hungarian 	& $36.016$			& $63.297$  		& $0.393$			\\ \hline
	Armenian  	& $9.814$			& $81.256$  		& $0.589$			\\ \hline
	Japanese  	& $0.196$			& $26.791$  		& $1.079$			\\ \hline
	Turkish   	& $2.355$			& $63.199$  		& $18.449$			\\ \hline
	Yoruba    	& $1.178$			& $48.577$  		& $36.997$			\\ \hline
	Basque    	& $0.393$			& $12.365$  		& $34.838$			\\ \hline
	Swahili   	& $14.328$ 			& $85.574$  		& $0.000$			\\ \hline
	Chinese   	& $1.865$ 			& $69.382$  		& $9.225$			\\ \hline
	\end{tabular}
	\caption{Percentage of female, male and neutral gender pronouns obtained for each language, averaged over all occupations detailed in Table \ref{tab:gender-neutral-languages}. Note that columns do not in general add up to $100\%$, as there is a fair amount of translated sentences for which we cannot obtain a gender pronoun.}
	\label{tab:gender-by-language}
	}
\end{table}

@@

% Table with our gender bias results for each adjective
\begin{table}[H]
\small{
	\centering
	\begin{tabular}{|c|c|c|c|c|}
	\hline
	Adjective 	& Female 	& Male 	& Neutral 	& Total 	\\ \hline
	\hline
	Happy		& $4$		& $3$	& $2$		& $11$		\\ \hline
	Wrong		& $0$		& $6$	& $4$		& $11$		\\ \hline
	Smart		& $2$		& $4$	& $3$		& $11$		\\ \hline
	Ashamed		& $3$		& $5$	& $1$		& $11$		\\ \hline
	Envious		& $2$		& $5$	& $1$		& $11$		\\ \hline
	Modest		& $1$		& $5$	& $3$		& $11$		\\ \hline
	Innocent	& $1$		& $7$	& $1$		& $11$		\\ \hline
	Sad			& $2$		& $5$	& $2$		& $11$		\\ \hline
	Afraid		& $1$		& $5$	& $0$		& $11$		\\ \hline
	Dumb		& $1$		& $5$	& $2$		& $11$		\\ \hline
	Kind		& $3$		& $3$	& $1$		& $11$		\\ \hline
	Loving		& $2$		& $5$	& $2$		& $11$		\\ \hline
	Arrogant	& $0$		& $8$	& $1$		& $11$		\\ \hline
	Helpless	& $2$		& $5$	& $2$		& $11$		\\ \hline
	Right		& $0$		& $6$	& $3$		& $11$		\\ \hline
	Brave		& $2$		& $6$	& $1$		& $11$		\\ \hline
	Proud		& $1$		& $7$	& $1$		& $11$		\\ \hline
	Cruel		& $1$		& $7$	& $2$		& $11$		\\ \hline
	Hateful		& $1$		& $4$	& $3$		& $11$		\\ \hline
	Guilty		& $0$		& $8$	& $0$		& $11$		\\ \hline
	Shy			& $5$		& $2$	& $2$		& $11$		\\ \hline
	Attractive	& $4$		& $1$	& $4$		& $11$		\\ \hline \hline
	Total 		& $38$		& $112$	& $41$		& -			\\ \hline 
	\end{tabular}
	\caption{Number of female, male and neutral pronominal genders in the translated sentences for each selected adjective.}
	\label{tab:gender-by-adjective}
	}
\end{table}
